---
title: "Sample size calculation for prevalence of sleep disturbance in senior Judo athletes in Canada"
subtitle: 'DOCUMENT: SAP-2022-036-TV-v01'
author: '**From:** Felipe Figueiredo **To:** Thiago Vivacqua'
date: "2022-12-03"
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    reference_docx: misc/style_SAP_en.docx
    toc: yes
# toc-title: "Sumário"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(scipen = 999)
library(pander)
library(knitr)
panderOptions('table.style', 'rmarkdown')
panderOptions('table.split.table', 160)

library(epiR)
library(tibble)

```

---

\newpage

**Sample size calculation for prevalence of sleep disturbance in senior Judo athletes in Canada**

**Document version**

```{r history}
Version <- c("01")
Changes <- c("Initial version")
history <- cbind(Version, Changes)
colnames(history) <- c("Version", "Alterations")
# pander(history, split.cells = 70)
kable(history)
```

```{r cleaning, include=FALSE}
# source("scripts/input.R", encoding = 'UTF-8')
```

---

# Abbreviations

- CI: confidence interval
- PSQI: Pittsburgh Sleep Quality Index

# Context

## Objectives

Calculate the minimum sample size to be able to detect the prevalence of sleep disturbance as measured by the Pittsburgh Sleep Quality Index in Judo athletes.

## Hypotheses

# Data

## Raw data

N/A

## Analytical dataset

N/A

# Study parameters

## Study design

## Inclusion and exclusion criteria

## Exposures

## Outcomes

**Specification of outcome measures** (Zarin, 2011):

1. (Domain)
2. (Specific measurement)
3. (Specific metric)
4. (Method of aggregation)

**Primary outcome**

## Covariates

N/A

# Statistical methods

## Statistical analyses

### Descriptive analyses

N/A

### Inferential analyses

N/A

### Statistical modeling

N/A

### Missing data

No missing data imputation will be performed.
<!-- All evaluations will be performed as complete case analyses. -->

## Significance and Confidence Intervals

All analyses will be performed using the significance level of 5%.
All significance hypothesis tests and confidence intervals computed will be
two-tailed.
<!-- left-tailed. -->
<!-- right-tailed. -->

## Study size and Power

```{r}
N <- 3899
Py <- .2
epsilon <- .5
se <- .896
sp <- .865
ss <- epi.sssimpleestb(N = N, Py = Py, epsilon = epsilon, se = se, sp = sp)

tibble(
  Parameter = c(
    "Target population size (chosen)",
    "True population prevalence (assumed)",
    "Tolerance (chosen)",
    "Diagnostic sensitivity (chosen)",
    "Diagnostic specificity (chosen)"
  ),
  Value = c(
    N |> as.character(),
    Py|> philsfmisc::format.pct(0),
    epsilon|> philsfmisc::format.pct(0),
    se|> philsfmisc::format.pct(),
    sp|> philsfmisc::format.pct()
  )
) |>
  kable(
    align = "c",
    caption = "**Table 1** Parameters for sample size calculation."
  )
```

- Sample size (calculated): `r ss`

## Statistical packages

This analysis will be performed using statistical software `R` version `r getRversion()`.

# Observations and limitations

**Recommended reporting guideline**

The adoption of the EQUATOR network (<http://www.equator-network.org/>) reporting guidelines have seen increasing adoption by scientific journals.
All observational studies are recommended to be reported following the STROBE guideline (von Elm et al, 2014).

# References

<!-- - **SAR-2022-036-TV-v01** -- [title] -->
<!-- - Cohen, J. (1988). Statistical power analysis for the behavioral sciences (2nd Ed.). New York: Routledge. -->
- Humphry RW, Cameron A, Gunn GJ. A practical approach to calculate sample size for herd prevalence surveys. Prev Vet Med. 2004 Oct 14;65(3-4):173-88. (<https://doi.org/10.1016/j.prevetmed.2004.07.003>).
- Zarin DA, et al. The ClinicalTrials.gov results database -- update and key issues. N Engl J Med 2011;364:852-60 (<https://doi.org/10.1056/NEJMsa1012065>).
- Gamble C, et al. Guidelines for the Content of Statistical Analysis Plans in Clinical Trials. JAMA. 2017;318(23):2337–2343 (<https://doi.org/10.1001/jama.2017.18556>).
- von Elm E, Altman DG, Egger M, Pocock SJ, Gøtzsche PC, Vandenbroucke JP; STROBE Initiative. The Strengthening the Reporting of Observational Studies in Epidemiology (STROBE) Statement: guidelines for reporting observational studies. Int J Surg. 2014 Dec;12(12):1495-9 (<https://doi.org/10.1016/j.ijsu.2014.07.013>).

# Appendix

This document was elaborated following recommendations on the structure for Statistical Analysis Plans (Gamble, 2017) for better transparency and clarity.

## Availability

All documents from this consultation were included in the consultant's Portfolio.

<!-- The client has requested that this analysis be kept confidential until a future date, determined by the client. -->
<!-- All documents from this consultation are therefore not published online and only the title and year of the analysis will be included in the consultant's Portfolio. -->
<!-- After the agreed date is reached, the documents will be released. -->

<!-- The client has requested that this analysis be kept confidential. -->
<!-- All documents from this consultation are therefore not published online and only the title and year of the analysis will be included in the consultant's Portfolio. -->

The portfolio is available at:

<https://philsf-biostat.github.io/SAR-2022-036-TV/>
